---
title: "Assignment 2 Question 1"
output: pdf_document
author: Sheen Thusoo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part d)

```{r}
data <- read.csv("EconomicMobility.csv")
```

``` {r}
correlation_coefficient <- function(pop_y, commute_y) {
  cor(pop_y, commute_y, method = "pearson", use='complete.obs')
}

delta = numeric(length(data$Population))
for (i in 1:length(data$Population)) {
  ## y[-i] removes the ith element from a vector
  delta[i] = correlation_coefficient(data$Population, data$Commute) - 
    correlation_coefficient(data$Population[-i], data$Commute[-i])
}
```


```{r}
plot(delta, main="Influence for Correlation Coefficient", ylab=bquote(Delta), pch=19, 
     col=adjustcolor("black", alpha = 0.2))
```


**Are there any influential points? If so, determine if there is anything interesting about them.**

```{r}
data[which(delta > 0.04),]
```


```{r}
delta[which(delta > 0.04)]
```

**Are there any influential points? If so, determine if there is anything interesting about them.**  
There is an influential point in the data which has an influence value that is much greater than other points (around 0.046). From the code chunk above, this point is in Los Angeles where the $Population$ variate is $16393360$ and the $Commute$ variate is $0.225$. However, compared to other cities in the dataset with a similar population value, the Commute variate is much lower; this is an outlier since it is much higher than what is expected. Thus, the correlation coefficient is much different than the others and the influence is much greater for this data point.


