---
title: 
- "STAT 341: Assignment 2 - Fall 2021"
author: "Instructor: Reza Ramezan"
date: 'Due: Monday, October 25 at 10:00am EDT'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r, include = FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Instructions

Your assignment must be submitted by the due date listed at the top of this document. The assignments deadlines are firm and the late submission penalty outlined in the course syllabus starts at 10:01am on Oct. 25. Note that this is an extended deadline relative to the tentative one in the course outline.

You must upload your solutions in the form of **one pdf file for each part of each question** by the deadline onto **Crowdmark** and/or **LEARN**. The response to one of the questions is in video format to be submitted on LEARN. Your instructor will NOT accommodate mistakes in submitting the pdf file of one question for another question. No assignment submission through email will be accepted. Note that your pdf solution file must have been generated by R Markdown. Additionally:

* For mathematical questions: your solutions must be produced by LaTeX (from within R Markdown), unless specifically mentioned that hand-written solutions are accepted where you can import a clear image of a hand-written solution. If such note is not included, screenshots and scanned/photographed handwritten solutions receive zero points.

* For computational questions: R code should always be included in your solution (via code chunks in R Markdown). If you don;t provide your R codes, you will receive zero points for the corresponding question.

* For interpretation questions: plain text (within R Markdown) is required. Text responses embedded as comments within code chunks will not be accepted.

* Alternative accommodations including, but not limited to, email submission and/or extensions due to RMarkdown breakdown and/or compilation to pdf will not be granted.

* The formatting requirement will be taken seriously from this assignment onward. Screenshots of your solutions and/or R codes, even if the original file is RMarkdown-generated, will receive 0 marks. Your submitted file to Crowdmark must be directly compiled by RMarkdown. 

* Question 3 of this assignment must be submitted in two places: on Crowdmark and on the Dropbox folder LEARN > Submit > Dropbox > Assignment 2 - Question 3

Organization and formatting is part of a full solution. Consequently, points will be deducted for solutions that are not organized and incomprehensible. A disorganized solution which is difficult to understand or find parts will not receive full marks. 

---

## Academic Integrity: 

* While you may discuss the questions with your classmates on Piazza, consulting another student's solution is prohibited, and submitted solutions may not be copied from any source. You may not talk to any other individual about the questions in this assignment. The instructor will hold online office hours during which he will answer clarification questions. You also have access to Piazza where you can ask questions.


* You may not use and/or search the internet (except for LEARN and Piazza) to answer the questions in this assignment. However, you may search the internet for R syntax.

* If a question which you would like to post on Piazza shares your solution, you must make it a private post.


* In short, you can treat this assignment like an open-book exam, where you are only allowed to use the course material provided to you during lectures and on Piazza and/or LEARN as well as books you may find at the library.


* Any violation of the the academic integrity regulations outlined here and in the course syllabus (make sure to read the course outline again!) will be counted as cheating and will be reported to the Dean's Office.

* The instructor reserves the right to conduct an online interview with you during which you will be asked questions about your solutions and the details of how you came to these responses. Should such an interview take place and you are unable to explain and defend your solutions, your grade for this assignment, and consequently, your course grade will be affected.

---

#### Question One - 14 Marks 
The `EconomicMobility` data-set (posted on LEARN) has information about 729 US communities to study the economic mobility across generations in the contemporary USA. The variables in the data-set are:

  - `Mobility`: The probability that a child born in 1980-1982 into the lowest quintile (20%) of household income will be in the top quintile (20%) at age 30. Individuals are assigned to the community they grew up in, not the one they were in as adults.
  
  - `Commute`: Fraction of workers with a commute of less than 15 minutes.
  
  - `Longitude / Latitude`: Geographic coordinate for the centre of the community.
  
  - `Name`: the name of principal city or town.
  
  - `State`: the state of the principal city or town of the community.
  
  - `Population`: The population of the community.


$\;$
$\;$

a) **[4 marks]** Construct a 2X1 figure containing two scatter-plots. The first plot is latitude (y) vs longitude (x) in which the points are shaded/coloured by population categories below.

    + Very low population: Population $\leq$ 100,000
    
    + Low population: Population $\in$ (100,000 $\,,\,$ 500,000]

    + High population: Population $\in$ (500,000 $\,,\,$ 5,000,000]

    + Very high population: Population $\geq$ 5,000,000
    
    The second scattor-plot is, again, lattitude (y) vs longitude (x) , but this time, points are shaded/coloured by commuting category as follows 

    + Very low commute: Commute $\leq$ 0.25
    
    + Low commute: Commute $\in$ (0.25 $\,,\,$ 0,5]
    
    + High commute: Commute $\in$ (0.5 $\,,\,$ 0.75]
    
    + Very high commute: Commute $\geq$ 0.75
    
    Comment on the patterns that you observe in each plot. Reading these plots, do you observe any relationship between the two variables `commute` and `population`?


$\;$
$\;$

b) **[2 marks]** We are interested in the relationship between the fraction of workers with a commute of less than 15 minutes (y) and the population of the communities (x). 

    + Write a function which takes in the pair $(\alpha_x\,,\,\alpha_y)$ and returns the power-transformed data (power transformation was explained in the lectures). 
    + Generate a 1x2 figure with two scatterplots. One with the orginal values of commute vs. population, and the other the scatterplot of the transformed data with $\alpha_x=\alpha_y=-1/2$. Make sure to use proper titles and axes labels in each plot
 

$\;$
$\;$

c) **[5 marks]** Using power-transformations, we would like to straighten the scatterplot in part (b). Suppose the strength of the linear relationship between two variables $X$ and $Y$ is measured by the attribute $L=1-\rho_{_{XY}}^2$ in which $\rho_{_{XY}}$ is the correlation coefficient between the two variables. Assuming $x=population$ and $y=commute$, find the power-transformed data $(T_{\alpha_x}(x),T_{\alpha_y}(y))$ which have the strongest linear relationship measured by $L$. For full marks, you should:

    + Create a function $L$ in `R` which takes in $\theta=(\alpha_x,\alpha_y)$ as input and outputs $1-corr^2\left(T_{\alpha_x}(x) , T_{\alpha_y}(y)\right)$.
    + Optimize the function $L$ with initial values $\alpha_x=\alpha_y=1$, and report optimum values of $\alpha_x$ and $\alpha_y$.
    + Generate the scatterplot of the original data as well as the scatterplot of the power-transformed data, and calculate the correlation coeeficient for each scatterplot
    
    While you have learned the gradient descent method for optimization, it requires deriving the gradient function. The built-in function `nlminb` can use different numerical optimization methods to find the minimum of a function. Read the help document of `nlminb` carefully, and use it for to solve your optimizaiton problem in this question.

$\;$
$\;$


d) **[3 marks]** Suppose $x_i$ is the population of the $i$th community and $y_i$ is the fraction of workers with a commute of less than 15 minutes in community $i$. Calculate the influence of observations $(x_i,y_i)$, $i=1,..,729$, on the correlation coefficient $\rho_{_{XY}}$ and generate the influence plot of the observations on the correlation coefficient. Are there any influential points? If so, determine if there is anything interesting about them.


---


#### Question Two - 24 Marks 
In STAT 231, maximum likelihood (ML) estimation was introduced. The ML is a parameter estimation method in which the likelihood function is maximized relative to the parameters of the distribution. This allows us to fit statistical distributions to the data.  In particular, based on independent observations $y_1,...,y_n$, the likelihood function of the parameter vector $\boldsymbol\theta$ is:
$$
L(\boldsymbol\theta;y_1,...,y_n) = f_{\boldsymbol\theta}(y_1)\times...\times f_{\boldsymbol\theta}(y_n)
$$
in which $f_{\boldsymbol\theta}(y_i)$ is the probability function of the variable $Y$ calculated at point $y_i$. For both theoretical and computational reasons, it is usually easier to maximize the log-likelihood function which is defined as the natural logarithm of the likelihood function, i.e., we usually maximize 
$$
\ell(\boldsymbol\theta;y_1,...,y_n) = \log\big(L(\boldsymbol\theta;y_1,...,y_n)\big)
$$
We believe that *Gamma* distribution with the following probability density function is an appropriate model for the variable `commute`.
$$
f(y) = \frac{\beta^\alpha y^{\alpha-1}}{\Gamma(\alpha)}\exp(-y\beta)
$$
where $\Gamma(\alpha)$ is the gamma function calculated at $\alpha$.

$\;$
$\;$

a) **[2 marks]** Derive the log-likelihood function of the parameter vector $\boldsymbol\theta=(\alpha,\beta)$. 


$\;$
$\;$


b) **[3 marks]** Construct a function called `CreateLogLikeFunction` that takes in a population or vector of observed values, and outputs a function of $\boldsymbol\theta=(\alpha,\beta)$ that calculates the log-likelihood function given $(\alpha,\beta)$ as input. Apply `CreateLogLikeFunction` on the variable `commute` form the economic mobility data and calculate the resulting function at $(\alpha,\beta) = (2,2)$. The function `gamma` calculates the gamma function in `R`.

$\;$
$\;$

c) **[3 marks]** Generate the 3D plot of the log-likelihood function in part (b) over the range $\alpha\in(0,100)$ and $\beta\in(0,100)$, as well as its heat map with the contour plot superimposed on the heat map. You may find the functions `outer`, `persp`, `image`, and `contour` useful for this task, but you do not have to use them.



$\;$
$\;$


d) **[5 marks]** Write a function which uses the gradient descent (or gradient ascent) with a line-search over the range $\lambda=0,0.01,0.02,...,5$ to find the maximum likelihood estimates of parameters $\alpha$ and $\beta$ given the variable `commute`. The function `digamma` in `R` calculates $\frac{d}{dx}\log(\Gamma(x))$.



$\;$
$\;$

e) **[4 marks]** Repeat the optimization in part (c) with  $\lambda=0,1,2,3,4,5$ and $\lambda=0.0001,0.0002,...,0.01$. In both cases keep the maximum number of iterations to 1000. What do you observe? Justify your observation.



$\;$
$\;$

f) **[4 marks]** Construct three 1x3 figures with the following plots:

    + the solution paths for $\alpha$ values versus the iteration number, $\beta$ values versus the iteration number, and the loglikelihood function $\ell(\alpha_i,\beta_i)$ versus the iteration number for the optimization problem in part (d), i.e. $\lambda=0,0.01,0.02,...,5$
    + a similar set of three plots above for the first optimization problem in (e), i.e. $\lambda=0,1,2,3,4,5$
    + a similar set of three plots above for the second optimization problem in (e), i.e. $\lambda=0.0001,0.0002,...,0.01$.
    



$\;$
$\;$

g) **[3 marks]** A probability histogram is a scaled histogram such that its total area is equal to one. The argument `probability=TRUE` in the function `hist` generates a probability histogram as opposed to a frequency histogram. Construct a 1X3 figure containing three plots. In each of the three panels plot a probability histogram of the variable `commute` and superimpose on it the probability density function of the Gamma distribution as follows.  

    + In the first panel replace $\alpha$ and $\beta$ in the pdf of the Gamma distribution with their estimates from part (d).
    + In the next two panels replace $\alpha$ and $\beta$ in the pdf of the Gamma distribution with their estimates from part (e).
    + Make sure to label the plots clearly.
    
    What do you observe? Justify your observation.



---

#### Question Three - 8 Marks

Gradient descent is an optimization algorithm commonly used in practice -- particularly in the machine learning community. For instance, [this job ad](https://www.indeed.com/viewjob?jk=a63a2c9ca658613c&tk=1f7jco166u45t801&from=serp&vjs=3) for an Applied Scientist at Amazon, and [this job ad](https://www.indeed.com/viewjob?jk=b47a3b5ee2d63549&tk=1f7jco166u45t801&from=serp&vjs=3) for a Principal Research Scientist on PepsiCo's Data Science & Analytics team, both explicitly indicate that the roles require experience with gradient descent. As such, in this question, you will answer two interview-style questions about gradient descent.

- Your response is limited to a **single page in total** (for part a and b) and should be comprised full sentences and possible formulas but no plots or figures. Imagine that you will be read out your response. 

- Additional to Crowdmark, you must also submit the solution to this question on LEARN > Submit > Dropbox where your response will be submitted to Turnitin (see the course outline for details on Turnitin). If you miss any of these platforms (Crowdmark or LEARN) you will automatically receive 0% on this question. For your response to this question to be marked, you must submit your 1-page pdf file on **both Crowdmark and LEARN**. 


**Rubric for each part**
```{r, echo=FALSE}
text_tbl <- data.frame(
Criteria = c("Writing", "Content", "Depth"),
Descriptor = c("Clarity & Grammar",
"Accuracy, Coverage and Terminology used", "Is the response at the appropriate level"),
Marks = c("/1", "/2", "/1")
)

knitr::kable(text_tbl, booktabs=T) 
```

$\;$

(a) **[4 Marks]** In your own words, describe the purpose of gradient descent and explain how it works. Aim your response at a level appropriate for that well-meaning, non-technical **family member** who at every holiday gathering asks you to explain what it is that you study. 

$\;$

(b) **[4 Marks]** In your own words, compare and contrast batch-sequential, batch-stochastic, and ordinary gradient descent, reflecting on the merits and drawbacks of each. Aim your response at a level appropriate for your **classmate** who read (and understood) the course notes up to the end of Section 2.3.2b, but who then fell asleep and didn't get to Section 2.3.2c.


---

#### Question Four - 4 Marks 

Create a short video (maximum 2 minutes) explaining the concept of influence, and the information that an influence plot conveys. You can create a presentation file from your R Markdown file (some instructions at https://rmarkdown.rstudio.com/lesson-11.html) or use any other software/package such as HTML, Beamer, LaTeX, or PowerPoint to make a presentation. While having a presentation file is not mandatory, it is highly encouraged and recommended as it will help you communicate your ideas more clearly. You may even include an example in your presentation file. To submit your video file, a separate video assignment has been created on LEARN/Bongo. You will be required to turn on your camera and share your screen. You may  record your response as many times as you want and only submit once you are happy with your recording. To avoid submission problems and lateness penalties, do not start recording just a few minutes prior to the deadline.

**Important Notes**:
  
  - To record and submit your video, go to LEARN > Content > Assignments > Assignment #2 > Link to A2-Q4.
  - Make sure you watch the instruction video once you click on the link for this question on LEARN to see what you exactly need to do.
  - You will need to turn on both your camera and microphone. Also, you will be required to show your WatCard at the beginning of your recording.
  - If your computer does not allow you to share your screen, you need to change the security settings on your computer.
  - Once you click submit, you cannot retake the assignment any more. All your retakes have to be done prior to submission.
  - You must click on submit once you are done with your recording. Failure to do so will result in no submission, hence a 0% mark on the question.
  - It is particularly important that you complete this question because future assignments may include questions which are worth significantly more marks, and this question will act as  practice/preparation for future assessments.
  - Additional help about completing this type of assignment on LEARN/Bongo is available at this link: https://bongolearn.zendesk.com/hc/en-us/articles/360005037594-How-to-Complete-Q-A

$\;$

$\;$

$\;$

$\;$