---
title: "Question 3"
author: "Sheen Thusoo"
output: pdf_document
---

## Write a function called drawBoxPlot that takes in a single input that is labelled df, an N×m data frame containing numeric data. This function should produce a single figure with m boxplots that share a common scale.

## - Use the $boxplot.stats$ to find the points for the figures.
## - Use the first 4 columns of the Iris data-set (length and width of sepals and petals) to demonstrate the $drawBoxPlot$ function.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, getting the first 4 columns of the Iris dataset.
``` {r}
iris_data <- read.csv("Iris.csv")
iris_df <- iris_data[,c("SepalLength","SepalWidth","PetalLength", "PetalWidth")]
```


Function:
``` {r}
drawBoxPlot <- function(df) {
  
  # Empty plot & legend
  plot(1:ncol(df), type="n", xlab = "Variate", ylab = "", main = "Boxplots", 
       xlim = c(0, ncol(df) + 0.5), ylim = c(0, max(df)), xaxt="n") 
  
  legend(4, 6, legend=c("Outliers"),
       col=c("black"), pch=(19), cex=0.5)
  
  axis(1, at=1:ncol(df), labels=colnames(df))
  
  for (i in 1:ncol(df)) {
    col_stats <- boxplot.stats(df[, colnames(df)[i]], do.out = TRUE)
    
    # Plot Outliers
    if (isFALSE(length(col_stats$out) == 0)) {
      for (outlier in 1: length(col_stats$out)) {
        points(i, col_stats$out[outlier], pch = 19, cex = 0.5)
      }
    }
    
    # Boxplot
    rect(i-0.2, as.numeric(unlist(col_stats$stats[2])), i+0.2, 
         as.numeric(unlist(col_stats$stats[3])), col = "#ece0f3") # rect from Q1 to median
    rect(i-0.2, as.numeric(unlist(col_stats$stats[3])), i+0.2, 
         as.numeric(unlist(col_stats$stats[4])), col= "#ece0f3") # rect from median to Q3
    
    # Vertical Lines
    lines(c(i,i), c(col_stats$stats[1], col_stats$stats[2]))
    lines(c(i,i), c(col_stats$stats[4], col_stats$stats[5]))

    # Whisker Lines
    lines(c(i-0.1,i+0.1), c(col_stats$stats[1], col_stats$stats[1]))
    lines(c(i-0.1,i+0.1), c(col_stats$stats[5], col_stats$stats[5]))
    
  }
}
```

Call the function:
``` {r}
drawBoxPlot(iris_df)
```

