---
title: "Assignment 1 Question 2"
author: "Sheen Thusoo"
output: pdf_document
---

## i. Using the SepalLength and PetalRatio, construct a 1×3 figure containing the two individual histograms and the scatterplot of the two variables.

Power Function provided in assignment.
``` {r}
powerfun <- function(x, alpha) {
  if(sum(x <= 0) > 1) stop("x must be positive")
  if (alpha == 0)
    log(x)
  else if (alpha > 0) {
    x^alpha
  } else -x^alpha
}
```


```{r}
iris_data <- read.csv("Iris.csv")
iris_data <- transform(iris_data, PetalRatio = PetalWidth / PetalLength)
par(mfrow=c(1,3)) 
plot( iris_data$SepalLength, iris_data$PetalRatio, pch = 19, 
      col=adjustcolor("firebrick", alpha = 0.5 ), 
      main = "Sepal Length vs. Petal Ratio", xlab = "Sepal Length (cm)", ylab = "Petal Ratio" )

hist(iris_data$SepalLength, breaks="FD", main=paste("Sepal Length Histogram"), 
     xlab = "Sepal Length (cm)")
hist(iris_data$SepalLength, breaks="FD", main=paste("Petal Ratio Histogram"), 
     xlab = "Petal Ratio")
```

## ii. What power (the values of $\alpha$) makes the distribution of SepalLength approximately symmetric? Plot a histogram of the transformed variable.

After trying various values of $\alpha$ using a for loop, including (-2, -1.5, -1, -0.2, -0.1, 0, 0.5, 1, 1.5), I found that the value of $\alpha$ that makes the distribution approximately symmetric is -0.2. The values above 1 made the graph left-skewed. I used Bump Rule #1 to find the most appropriate value. Because the bump was on lower values when $\alpha$ was greater than 1, I moved the power lower on the ladder. I tried values like -3, and -2 but noticed that they produced a bump on the higher values, so I needed to move the power up on the ladder. After trying many other values that are shown below, I noticed that a value of -0.2 made the histogram symmetric.

```{r}
par(mfrow=c(2,3)) 
a = c(-3, -2, -0.2, 0, 1, 1.5)
for (i in 1:length(a) ) hist( powerfun(iris_data$SepalLength + 1, a[i]), breaks="FD", 
                               main=bquote(alpha==.(a[i])), xlab = "Sepal Length (cm)", a[i])
```

The histogram for the chosen $\alpha$ value is shown below. 
```{r}
a = -0.2
hist( powerfun(iris_data$SepalLength + 1, a), breaks="FD", 
                               main=bquote(alpha==.(a)), xlab = "Sepal Length (cm)", a)
```

The value $\alpha = -0.2$ makes the distribution of Sepal Length approximately symmetric.

## iii. What power (the values of $\alpha$) makes the distribution of PetalRatio approximately symmetric? Plot a histogram of the transformed variable.

After trying various values of $\alpha$ using a for loop, including (-0.1, 0, 0.5, 0.6, 1, 1.5), I found that the value of $\alpha$ that makes the distribution approximately symmetric is 3.5. I used Bump Rule #1 to find the most appropriate value. Because the bump was on higher values when $\alpha$ was lower than 1, I moved the power higher on the ladder. I tried values like 2.5 and 3 but noticed that they produced a bump on the lower values, so I needed to move the power down on the ladder. After trying many other values that are shown below, I noticed that a value of -0.2 made the histogram symmetric.


```{r}
par(mfrow=c(2,3))
a1 = c(-2, -1, 0, 1, 2, 3.5)
for (i in 1:length(a1) ) hist( powerfun(iris_data$PetalRatio + 1, a1[i]), breaks="FD", 
                               xlab = "Petal Ratio", main=bquote(alpha==.(a1[i])), a1[i])
```
The histogram for the chosen $\alpha$ value is shown below. 
```{r}
a = 3.5
hist( powerfun(iris_data$PetalRatio + 1, a), breaks="FD", 
                               main=bquote(alpha==.(a)), xlab = "Petal Ratio", a)
```

The value $\alpha = 3.5$ makes the distribution of Petal Ratio approximately symmetric.

## iv. What pair of powers (the values of $\alpha$) makes the scatter plot of SepalLength and PetalRatio approximately linear? Plot a scatter plot of the transformed variables.

The pairs of values that make the scatterplot approximately linear are $\alpha_x = -0.2$ and $\alpha_y = 3.5$. I used bump rule #2. I found that these values worked best in making the scatterplot similar. When zooming out on the x-axis slightly, the plot looks more linear, as shown in the second figure below.

``` {r}
plot( powerfun(iris_data$SepalLength + 1, -0.2), powerfun(iris_data$PetalRatio + 1, 3.5), 
      pch = 19, col=adjustcolor("firebrick", alpha = 0.5 ), xlab = "Sepal Length (cm)", 
      ylab = "Petal Ratio", main = "Sepal Length vs. Petal Ratio")
abline( lm( I(powerfun(iris_data$PetalRatio + 1, 3.5)) ~ 
              I(powerfun(iris_data$SepalLength + 1, -0.2))), col="blue" )
```

``` {r}
plot( powerfun(iris_data$SepalLength + 1, -0.2), powerfun(iris_data$PetalRatio + 1, 3.5), 
      pch = 19, col=adjustcolor("firebrick", alpha = 0.5 ), xlim= c(-0.8, -0.5), 
      xlab = "Sepal Length (cm)", ylab = "Petal Ratio", main = "Sepal Length vs. Petal Ratio")
abline( lm( I(powerfun(iris_data$PetalRatio + 1, 3.5)) ~ 
              I(powerfun(iris_data$SepalLength + 1, -0.2))), col="blue" )
```


